---
title: "STAT 240 Discussion 9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
source("../../scripts/viridis.R")
source("../../scripts/ggprob.R")
```

\newcommand{\E}{\mathsf{E}}
\newcommand{\Var}{\mathsf{Var}}
\renewcommand{\prob}{\mathsf{P}}

## Problems

The questions in this discussion assignment use the chimpanzee data set.
In answering these questions, you may assume the following in models for the data:

1. all trials are independent
2. trials made with a partner have the same chance of being pro-social, even if the identities of the actors or partners differ.

### Problem 1

#### A

Using the chimpanzee data,
combine all of the trials made without a partner.
Count the number of pro-social and selfish choices made.

```{r}
chimp = read.csv("../../data/chimpanzee.csv")

chimp_c = chimp %>%
  filter(partner != "none") %>%
  summarise(prosocial_c = sum(prosocial), selfish_c = sum(selfish))
```

#### B

Write a model for the number of pro-social choices.

> It is a binomial model with n = 610, p = 0.588

#### C

Use the Agresti-Coull method to construct a 95% confidence interval for the probability that a chimpanzee makes the pro-social choice in experimental conditions when there is no partner present.

```{r}
binom_se =  function(n, p){
  return ( sqrt( p*(1-p)/n) )
}
binom_ci = function(est, se, conf=0.95){
  z = qnorm(1 - (1 - conf)/2)
  me = z * se
  ci = est + c(-1,1)*me
  return(ci)
}

n = chimp_c$prosocial_c + chimp_c$selfish_c
p = (chimp_c$prosocial_c + 2) / (n + 4)
se = binom_se(n+4, p)

ci_agresti = binom_ci(p, se)
round(ci_agresti, 3)
```

#### D

Interpret this confidence interval in the context of the experimental setting.

> We are 95% certain that the probability that a chimp with no partner makes a pro-social choice is between p = 0.549 and 0.627

### Problem 2

Using the same data as in the previous problem,
do the following.

#### A

State hypotheses for a two-sided hypothesis test that the probability that a chimpanzee makes the pro-social choice over the long run is equal to 0.5 versus the alternative that this probability is not 0.5.

> H0 => p = 0.5
  H1 => p != 0.5

#### B

Pick a test statistic, state the sampling distribution of this statistic if the null hypothesis is true,
and compute a p-value for the test.

```{r}

```

#### C

Interpret these results in the context of the experiment following the examples from lecture notes.

### 3

Use functions from the file `ggprob.R` to visualize the calculation of the p-value by graphing an appropriate probability distribution and using differential coloring to indicate what in the plot corresponds to the p-value.
Include a title and meaningful axis labels.

```{r}

```

