---
author: "Abhinandan Saha"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE, error = TRUE, fig.height = 3)
library(tidyverse)
library(kableExtra)
source("../../scripts/viridis.R")
source("../../scripts/ggprob.R")
```

\newcommand{\E}{\mathsf{E}}
\newcommand{\Var}{\mathsf{Var}}
\renewcommand{\prob}{\mathsf{P}}


## Assignment 7

#### Due Sunday, April 2, 11:59 PM CT

### Preliminaries

- Directories
    - COURSE/homework/
    - COURSE/homework/hw07/
    - COURSE/data/
    - COURSE/scripts/
- Files
  - COURSE/homework/hw07/hw07.Rmd
  - COURSE/scripts/viridis.R
  - COURSE/scripts/ggprob.R

### Aims

- Practice probability and the binomial distribution

### Problems

  1. A discrete random variable $X$ has possible values 0, 1, 2, 3, 4, 5, 6, 7, 8 with the following partial probability distribution.
The missing probabilities $\prob(X=7)$ and $\prob(X=8)$ are equal to each other.

```{r, echo = FALSE}
set.seed(20221021)
x = 0:8
p = rnorm(7,5,2.1)
p = sort(round(p / sum(p) * 0.84, 2))
p1 = tibble(x = x, p = c(p, rep(NA,2)))

p1_tab = p1 %>% 
  mutate(x = str_c(x)) %>% 
  pivot_wider(names_from = x, values_from = p) %>% 
  mutate(x = "P(X=x)", .before = `0`) 

p1_tab %>% 
  kable() %>% 
  kable_styling(position = "center", full_width = FALSE)
```

- Put the values `x = 0:8` and the tabled probabilities `p` into a data frame (use `tibble()`).

```{r}
dist1 = tibble(x = x, p = c(p, rep(NA,2)))
```

- Write code to calculate the missing values for $\prob(X=7) = \prob(X = 8)$ and replace the missing values with these calculated values into your data frame.

```{r}
req = 1 - sum(dist1$p, na.rm = TRUE)
dist1[8,2] = req/2
dist1[9,2] = req/2
```

- Verify that the sum of all probabilities is 1.

```{r}
sum(dist1$p)
```

- Calculate the mean, variance, and standard deviation of this distribution.

```{r}
m = sum(dist1$x * dist1$p)
v = sum(dist1$x * dist1$p *(1 - dist1$p))
s = sqrt(v)
m
v
s
```


  2. Suppose you have a random variable $X \sim \text{Binomial}(120, 0.2)$.

- 2a. Calculate and report the mean and standard deviation of the distribution.

```{r}
m1 = 120*0.2
sd1 = sqrt((120*0.2)*0.8)
m1
sd1
```


- 2b. Calculate and report the probability that the random variable is exactly equal to 20, $\prob(X = 20)$.

```{r}
r1 = dbinom(20, 120, 0.2)
r1
```


- 2c. Calculate and report the probability that $X$ equals 20 or more, $\prob(X \ge 20)$.

```{r}
r2 = 1 - pbinom(19, 120, 0.2)
r2
```


- 2d. Draw a graph of the binomial distribution with vertical lines at the mean (red and dashed), and 1, 2, and 3 standard deviations above and below the mean (red and dotted).


```{r}
gbinom(120, 0.2, scale = TRUE) +
  geom_vline(xintercept = m1, color = "red",linetype = "dashed") +
  geom_vline(xintercept = m1 + c(-1,1)*sd1,
             color = "red", linetype = "dashed") +
  geom_vline(xintercept = m1 + c(-2,2)*sd1,
             color = "red", linetype = "dashed") +
  geom_vline(xintercept = m1 + c(-3,3)*sd1,
             color = "red", linetype = "dashed")
```


  3. What is the probability that $X$ from Problem 2 is within one, two, and three standard deviations of the mean? Round each probability to four decimal places.

```{r}
p1 = round(pbinom(m1 + sd1, 120, 0.2) - pbinom(m1 - sd1, 120, 0.2), 4)
p2 =  round(pbinom(m1 + (2*sd1), 120, 0.2) - pbinom(m1 - (2*sd1), 120, 0.2),4)
p3 =  round(pbinom(m1 + (3*sd1), 120, 0.2) - pbinom(m1 - (3*sd1), 120, 0.2),4)
p1
p2
p3
```

> One standard deviation of mean = 0.696
Two standard deviation of mean = 0.9486
Three standard deviation of mean = 0.9979

  4. The following code makes a graph of the cumulative distribution function for the $\text{Binomial}(7, 0.3)$ distribution, that is $F(x) = \prob(X \le x)$. This is the function which is calculated by the base R function `pbinom()`.

- Modify the code to add a horizontal red dotted line at 0.4.


```{r}
prob4 = tibble(
  x = seq(-1,8,1),
  p = dbinom(x, 7, 0.3),
  cdf = pbinom(x, 7, 0.3))

ggplot(prob4, aes(x = x, y = cdf )) +
  geom_step(color = "blue") +
  geom_hline(yintercept = 0.4, linetype = "dashed", color = "red") +
  scale_x_continuous(breaks = 0:7) +
  xlab("x") +
  ylab("Probability") +
  ggtitle("Binomial(7, 0.3) distribution CDF") +
  theme_minimal()
```

- Calculate the 0.4 quantile of the $\text{Binomial}(7, 0.3)$

```{r}
qbinom(0.4, 7, 0.3)
```

- Explain how the value of the quantile is related to the red dotted line which you added to the plot.

> The red dotted line represents the 0.4 quantile or the 40th percentile. It coincides with x=2, implying that x=2 is the 40th percentile of this distribution.

- Use the graph to determine the 0.75 quantile of this distribution.
Then verify your observation with a calculation.

```{r}
3

qbinom(0.75, 7, 0.3)
```

> The 0.75 quantile of this distribution is 3.



  5. In a large population of likely voters, the proportion of voters who plan to vote for candidate A is $p = 0.52$. In a random sample of 600 likely voters, what is the probability that the proportion of voters who plan to vote for candidate A is less than 50% (the number of such voters is less than 300)?
  
```{r}
pbinom(299, 600, 0.52)
```

> In a random sample of 600 likely voters, the probability that the proportion of voters who plan to vote for candidate A is less than 50% (the number of such voters is less than 300) is 15.35275%.


  6. Basketball player 1 makes 80% of their free throws and basketball players 2 makes 70% of their free throws. Assume all of their free throws attempts are independent. If each player takes ten free throws and $X$ is the total number of free throws they make, does $X$ have a binomial distribution? If yes, specify the parameters $n$ and $p$. If not, explain why not.
  
> It is a binomial distribution, since the number of trials is fixed to 10 by the problem, the trials are independednt of each other and does not change from trial to trial and each trial may end in either a success or failure - making the shot or not. The parameters are n = 10, p = 0.8 for player1 and 0.7 for player2.




  7. What is the smallest number of coin tosses $n$ of a fair coin so that the chance of getting 50 or more heads exceeds 90%?  (You may need to do some trial and error. Alternatively, think of creating a data frame with a column `n` and a column for the probability and using this to find the answer.)

```{r}
1-pbinom(49, 113, 0.5)
```

> n = 113



  8. The random variable $X$ has the $\text{Binomial}(100, 0.2)$ distribution.
  
- Find an integer $a$ so that $P(X \le a) \ge 0.5$ and $P(X \ge a) \ge 0.5$.
- Show the values of $a$, $P(X \le a)$, and $P(X \ge a)$.

```{r}
x = 0:100
df = data.frame(x, p=0.2)
df %>%
  mutate(le = pbinom(x, 100, 0.2), re = 1-pbinom(x-1, 100, 0.2, )) %>%
  filter(le>= 0.5 & re>=0.5)
```

> a = 20, P(X<=a) = 0.5594616, P(X>=a) = 0.5398386
  


